---
title: "Visualize GEMMA outputs"
author: "Miles Roberts"
date: "2021-03-24"
output: html_document
---

## Goals
* Identify significant hits from multivariate and univariate GWAS runs in GEMMA
* Visualize hits with manhattan plots

```{r}
rm(list = ls())
library(qqman)
library(data.table)
library(ggplot2)
library(R.utils)
wd = "C:\\Users\\miles\\OneDrive\\Education\\MichiganStateUniversity\\Josephs_Lab_Projects\\pleiotropyGWAS\\01_arabidopsis_trait_correlations\\attempt7"
setwd(paste(wd,"\\output", sep = ""))

pvalues = fread("mvKmerGWASresults_FT16_RGR_gxe.pvalues.txt.gz")
head(pvalues)
pvalues$p_wald_bh = p.adjust(pvalues$p_wald, method = "BH")
head(pvalues)

#What is the bh-corrected cutoff for significance?
max(pvalues$p_wald[which(pvalues$p_wald_bh < 0.05)])
```


```{r}

#Load SNP positions on chromosomes
wd = "C:\\Users\\miles\\OneDrive\\Education\\MichiganStateUniversity\\Josephs_Lab_Projects\\pleiotropyGWAS\\01_arabidopsis_trait_correlations\\attempt7"
setwd(wd)
pos = fread("positions.txt", sep = "\t", header = T)

#Write function to convert SNP identifiers to chromosomal locations
id_to_chrom = function(x){
	all_snpids = as.numeric(gsub("rs","",x))
	all_chrom = all_snpids

	all_chrom[which(0 <= all_snpids & all_snpids <= 2597734)] = 1

	all_chrom[which(2597735 <= all_snpids & all_snpids <= 4466529)] = 2

	all_chrom[which(4466530 <= all_snpids & all_snpids <= 6660781)] = 3

	all_chrom[which(6660782 <= all_snpids & all_snpids <= 8427785)] = 4

	all_chrom[which(8427786 <= all_snpids & all_snpids <= 10709465)] = 5

	return(all_chrom)

}

#Write function to format results for qqman package plots (manhattan plot, qqplot)
gemma_to_qqman = function(fitsnps, pos){

	#Extract snp ids, convert to chromosome information
	all_snpids = as.numeric(gsub("rs","",fitsnps$rs))
	all_chrom = id_to_chrom(fitsnps$rs)

	#Extract position information for snps
	all_pos = as.data.frame(pos[which(pos$V1 %in% all_snpids), 2])
	names(all_pos) = NULL

	#Build dataframe for genome wide manhattan plot
	my_results = data.frame(SNP = fitsnps$rs, CHR = all_chrom, BP = all_pos, P = fitsnps$p_wald)
	return(my_results)
}
```


## mvGWAS with GxE
```{r}
setwd(paste(wd,"\\output", sep = ""))
mysnps = fread("mvKmerGWASresults_FT16_RGR_gxe_GoodCol.assoc.txt.gz", select = c(1,7))
#head(mysnps)
#summary(mysnps$p_wald)

mysnps$p_wald_bf = p.adjust(mysnps$p_wald, method = "bonferroni")
mysnps$p_wald_bh = p.adjust(mysnps$p_wald, method = "BH")

#How many SNPs pass thresholds
mysnps[which(mysnps$p_wald_bf < 0.05),]
mysnps[which(mysnps$p_wald_bh < 0.05),]
sigsnps = mysnps[which(mysnps$p_wald_bh < 0.05),]
sigsnpsGxE = mysnps[which(mysnps$p_wald_bh < 0.05),]

#Correlate effect sizes of SNPs
plot(sigsnps$beta_1, sigsnps$beta_2)
cor(sigsnps$beta_1, sigsnps$beta_2)

#Extract results for qqman plots
mvGWASresults = gemma_to_qqman(mysnps, pos)
sigsnpLoc = mvGWASresults[which(mvGWASresults$SNP %in% sigsnps$rs),]

sigsnpLoc[order(sigsnpLoc$CHR, sigsnpLoc$BP),]
#Plot results
manhattan(mvGWASresults, suggestiveline = FALSE, genomewideline = -log10(0.05/nrow(mvGWASresults)), col = c("black", "deepskyblue"))

qq(mvGWASresults$P)

```

## mvGWAS without GxE
```{r}
setwd(paste(wd,"\\output", sep = ""))
mysnps = fread("mvGWASresults_FT16_RGR_NoGxe.assoc.txt")
mvGWASsnps = mysnps
#head(mysnps)
#summary(mysnps$p_wald)

mysnps$p_wald_bf = p.adjust(mysnps$p_wald, method = "bonferroni")
mysnps$p_wald_bh = p.adjust(mysnps$p_wald, method = "BH")

#How many SNPs pass thresholds
mysnps[which(mysnps$p_wald_bf < 0.05),]
mysnps[which(mysnps$p_wald_bh < 0.05),]
sigsnps = mysnps[which(mysnps$p_wald_bh < 0.05),]
sigsnpsNoGxE = mysnps[which(mysnps$p_wald_bh < 0.05),]

#Correlate effect sizes of SNPs
plot(sigsnps$beta_1, sigsnps$beta_2)
cor(sigsnps$beta_1, sigsnps$beta_2)

#Extract results for qqman plots
mvGWASresults = gemma_to_qqman(mysnps, pos)
sigsnpLoc = mvGWASresults[which(mvGWASresults$SNP %in% sigsnps$rs),]

sigsnpLoc[order(sigsnpLoc$CHR, sigsnpLoc$BP),]
#Plot results
manhattan(mvGWASresults, suggestiveline = FALSE, genomewideline = -log10(0.05/nrow(mvGWASresults)), col = c("black", "deepskyblue"))

qq(mvGWASresults$P)
```

## Compare mvGWAS results with and without GxE
```{r}
sigsnpsGxE
sigsnpsNoGxE
any(sigsnpsGxE$rs %in% sigsnpsNoGxE$rs)
```

## uvGWAS results
### FT Without GxE
```{r}
setwd(paste(wd,"\\output", sep = ""))
mysnps = fread("uvGWASresults_FT16_NoGxe.assoc.txt")
ft16snps = mysnps

head(mysnps)
summary(mysnps$p_wald)

mysnps$p_wald_bf = p.adjust(mysnps$p_wald, method = "bonferroni")
mysnps$p_wald_bh = p.adjust(mysnps$p_wald, method = "BH")

#How many SNPs pass thresholds
mysnps[which(mysnps$p_wald_bf < 0.05),] #bonferroni threshold
mysnps[which(mysnps$p_wald_bh < 0.05),] #FDR threshold
sigsnps = mysnps[which(mysnps$p_wald_bh < 0.05),]

#Extract results for qqman plots
uvGWASresults = gemma_to_qqman(mysnps, pos)

#Plot results
manhattan(uvGWASresults, suggestiveline = FALSE, genomewideline = -log10(0.05/nrow(uvGWASresults)), col = c("black", "deepskyblue"))

qq(uvGWASresults$P)
```

### RGR Without GxE
```{r}
setwd(paste(wd,"\\output", sep = ""))
mysnps = fread("uvGWASresults_RGR_NoGxe.assoc.txt")
rgrsnps = mysnps

#head(mysnps)
#summary(mysnps$p_wald)

mysnps$p_wald_bf = p.adjust(mysnps$p_wald, method = "bonferroni")
mysnps$p_wald_bh = p.adjust(mysnps$p_wald, method = "BH")

#How many SNPs pass thresholds
mysnps[which(mysnps$p_wald_bf < 0.05),]
mysnps[which(mysnps$p_wald_bh < 0.05),]

#Extract results for qqman plots
uvGWASresults = gemma_to_qqman(mysnps, pos)

#Plot results
manhattan(uvGWASresults, suggestiveline = FALSE, genomewideline = -log10(0.05/nrow(uvGWASresults)), col = c("black", "deepskyblue"))

qq(uvGWASresults$P)
```

## Compare mvGWAS and uvGWAS
```{r}
mvuvComp = merge(mvGWASsnps, ft16snps, by = "rs")
mvuvComp = merge(mvuvComp, rgrsnps, by = "rs")

cor(mvuvComp$beta_1, mvuvComp$beta.x)
cor(mvuvComp$beta_2, mvuvComp$beta.y)

cor(mvuvComp$beta_2, mvuvComp$beta.x)
cor(mvuvComp$beta_1, mvuvComp$beta.y)

ggplot(mvuvComp, aes(beta_1, beta.x)) +
  geom_point() +
  theme_classic()

ggplot(mvuvComp, aes(beta_2, beta.y)) +
  geom_point() +
  theme_classic()
```

## Compare dataset with imputed phenotypes to dataset without imputed phenotypes
```{r}
setwd(paste(wd,"\\output", sep = ""))
imputedPheno = read.table("phenos_FT16_RGR_imputed.prdt.txt", sep = "\t")

setwd(wd)
originalPheno = read.table("phenos_FT16_RGR.txt", sep = "\t")

cor(originalPheno$V1, originalPheno$V2, use = "pairwise.complete.obs")
cor(imputedPheno$V1, imputedPheno$V2, use = "pairwise.complete.obs")

plot(originalPheno$V1, originalPheno$V2)
plot(imputedPheno$V1, imputedPheno$V2)
```

## PCGWAS without GxE
```{r}
setwd(paste(wd,"\\output", sep = ""))
mysnps = fread("pcGWASresults_FT16_RGR_NoGxe.assoc.txt")
#head(mysnps)
#summary(mysnps$p_wald)

mysnps$p_wald_bf = p.adjust(mysnps$p_wald, method = "bonferroni")
mysnps$p_wald_bh = p.adjust(mysnps$p_wald, method = "BH")

#How many SNPs pass thresholds
mysnps[which(mysnps$p_wald_bf < 0.05),]
mysnps[which(mysnps$p_wald_bh < 0.05),]

#Extract results for qqman plots
pcGWASresults = gemma_to_qqman(mysnps, pos)

#Plot results
manhattan(pcGWASresults, suggestiveline = FALSE, genomewideline = -log10(0.05/nrow(pcGWASresults)), col = c("black", "deepskyblue"))

qq(pcGWASresults$P)
```
